---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# palap

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of palap is to provide "reflective" colour palettes. Broadly speaking, there are three categories of colour palette:

1. **Qualitative**: categorical variables
2. **Sequential**: low to high numeric values
3. **Diverging**: negative to positive values

However,a fourth colour palette is needed, to cover cases where things diverge, but maintain some constant value/meaning as they get further from the center point. The proposed name for this is a "reflective" colour palette.

At the moment this package exists as a proof of concept, and a place to explore this idea. It is currently drawing heavily from [scico](https://github.com/thomasp85/scico).

## Installation

You can install the released version of palap from github with:

``` r
# install.packages("remotes")
remotes::install_github("njtierney/palap")
```

# Example

```{r example}
library(palap)
library(prismatic)
library(paletteer)

regular <- paletteer_d(package = "RColorBrewer", palette = "PuBu", n = 8)

# use the colour and plot functions from prismatic
plot(colour(regular))
```


```{r example-palap}
# now show this with palap
sym_pal <- palap(8)

plot(colour(sym_pal))
```


```{r example-reverse}
plot(colour(palap(8, direction = -1)))
```

In the future there may be a magic function that allows for creation of "reflective" colour palettes - perhaps similary to [paletti](https://github.com/EdwinTh/paletti).


# Example from `brolgar`

```{r}
library(dplyr)
library(brolgar)
heights_near <- key_slope(heights,
          height_cm ~ year) %>%
  keys_near(key = country,
            var = .slope_year) %>%
  left_join(heights, by = "country")

library(ggplot2)
p <- ggplot(heights_near,
       aes(x = year,
           y = height_cm,
           group = country,
           colour = stat)) +
  geom_line()

p

# lhs
p + scale_colour_palap_d(begin = 0,
                         end = 0.5)

# rhs?
p + scale_colour_palap_d(begin = 0.5,
                         end = 1)

p + scale_colour_palap_d(direction = -1)

```

