---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# palap

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of palap is to provide "symmetric" colour palettes. Broadly speaking, there are three categories of colour palette:

1\. **Qualitative**: categorical variables

2\. **Sequential**: low to high numeric values

3\. **Diverging**: negative to positive values

However, a fourth colour palette is needed, to cover cases where things diverge, but maintain some constant value/meaning as they get further from the center point. The proposed name for this is a "symmetric" colour palette:

4\. **Symmetric**: Diverging values with equivalent meaning as they get further from the diverging point.

At the moment this package exists as a proof of concept, and a place to explore this idea. It is currently drawing heavily from [scico](https://github.com/thomasp85/scico).

## Installation

You can install the released version of `palap` from github with:

``` r
# install.packages("remotes")
remotes::install_github("njtierney/palap")
```

# Example

```{r example}
library(palap)
library(prismatic)
library(paletteer)

regular <- paletteer_d(package = "RColorBrewer", palette = "PuBu", n = 8)

# use the colour and plot functions from prismatic
plot(colour(regular))
```


```{r example-palap}
# now show this with palap
sym_pal <- palap(8)

plot(colour(sym_pal))
```


```{r example-reverse}
plot(colour(palap(8, direction = -1)))
```

# Future work

It will be possible to generate your own symmetric colour palette from a given vector of colours.

- perhaps similary to [paletti](https://github.com/EdwinTh/paletti).


# Example from `brolgar`

```{r}
library(dplyr)
library(brolgar)
heights_near <- key_slope(heights,
          height_cm ~ year) %>%
  keys_near(key = country,
            var = .slope_year) %>%
  left_join(heights, by = "country")

library(ggplot2)
p <- ggplot(heights_near,
       aes(x = year,
           y = height_cm,
           group = country,
           colour = stat)) +
  geom_line()

p

# lhs
p + scale_colour_palap_d(begin = 0,
                         end = 0.5)

# rhs?
p + scale_colour_palap_d(begin = 0.5,
                         end = 1)

p + scale_colour_palap_d(direction = -1)

```

# Acknowledgements

Thank you to [Di Cook](http://dicook.org/) for pointing out the need for this kind of colour palette, it is one of those things that once you notice, you cannot unsee, and that's so cool. Thanks also to [Alison Hill](https://alison.rbind.io/) for pointing out that I could call this "symmetric" instead of "reflective" - to me this is much clearer.
